Let's talk about Thurstone \& Co.: An information-theoretical model for comparative judgments, and its statistical translation

### 2024-11-15 _tex/index.tex
Built site for gh-pages

Surprisingly, despite its reliance on the largest number of simplifying
assumptions \citep[pp.~253]{Bramley_2008}, Case V remains the most
widely used case in the CJ literature. This popularity is largely due to
its simplified statistical representation in the Bradley-Terry-Luce
(BTL) model \citep{Bradley_et_al_1952, Luce_1959}. The BTL model mirrors
Case V's assumptions, with one key difference: while Case V assumes a
Normal distribution for the discriminal processes, the BTL model uses
the more mathematically tractable logistic distribution
\citep[pp.~254]{Bramley_2008} (see Figure~\ref{fig-thurstone_cases}).
This substitution has little effect on the model's estimation or
interpretation, as the Normal and logistic distributions differ by a
scaling factor of approximately \(1.7\)
\citep[pp.~16]{vanderLinden_et_al_2017_I} (refer to
Figure~\ref{fig-logistic_vs_normal}).

@@

However, Case V (and its BTL model counterpart) was originally developed
to provide a ``rather coarse scaling'' of traits
\citep[pp.~269]{Thurstone_1927b}, prioritizing statistical simplicity
over precision in trait measurement. As a result, its assumptions may
not be suitable for applications beyond the psychophysical contexts,
which typically involve simple, homogeneous stimuli. Thurstone himself
cautioned that its use ``should not be made without (an) experimental
test'' \citep[pp.~270]{Thurstone_1927b}, further acknowledging that some
assumptions could become problematic with less homogeneous stimuli, such
as handwriting or English compositions \citep[pp.~374]{Thurstone_1927a}.
Consequently, given that current CJ applications often involve more
complex, less homogeneous stimuli, two assumptions of Case V may not
always hold theoretical or in practice: the assumption of zero
correlation and equal dispersion between stimuli.

>>>>

Surprisingly, despite its reliance on the largest number of simplifying
assumptions \citep[pp.~253]{Bramley_2008}, Case V remains the most
widely used case in the CJ literature. This popularity is largely due to
its simplified statistical representation in the Bradley-Terry-Luce
(BTL) model \citep{Bradley_et_al_1952, Luce_1959}. The BTL model mirrors
Case V's assumptions, with one key difference: while Case V assumes a
Normal distribution for the discriminal processes, the BTL model uses
the more mathematically tractable Logistic distribution
\citep[pp.~254]{Bramley_2008} (see Figure~\ref{fig-thurstone_cases}).
This substitution has little effect on the model's estimation or
interpretation, as the Normal and Logistic distributions differ by a
scaling factor of approximately \(1.7\)
\citep[pp.~16]{vanderLinden_et_al_2017_I} (refer to
Figure~\ref{fig-logistic_vs_normal}).

@@

However, Case V (and its BTL model counterpart) was originally developed
to provide a ``rather coarse scaling'' of traits
\citep[pp.~269]{Thurstone_1927b}, prioritizing statistical simplicity
over precision in trait measurement. As a result, its assumptions may
not be suitable for applications beyond the psychophysical contexts,
which typically involve simple, homogeneous stimuli. Thurstone himself
cautioned that its use ``should not be made without (an) experimental
test'' \citep[pp.~270]{Thurstone_1927b}, further acknowledging that some
assumptions could become problematic with less homogeneous stimuli, such
as handwriting or English compositions \citep[pp.~374]{Thurstone_1927a}.
Consequently, given that current CJ applications often involve more
complex, less homogeneous stimuli, two assumptions of Case V may not
always hold theoretically or in practice: the assumption of zero
correlation and equal dispersion between stimuli.
